---
title: "Baseline & Stress: Heart Rate"
author: "Brandon L. Kramer, et al."
output: html_document
---

This study, entited *Baseline & Stress*, was designed (1) to determine the best “resting state” method to ensure that participants are at their baseline levels of heart rate, respiratory activity, and blood pressure before starting psychophysiological experiments and (2) to examine how individuals respond to well-estbalish laboratory stressor called the Stroop task. For a more comprehensive overview, see the Study Overview and IRB proposals in the docs folder. 

The first step is concatenating all of the data and reducing the dimensions to carry out a comparison of best baselines tests and to examine the physiological reactivity of participants in the study. Given the complex nature of psychophysiological research 

```{r setup, include=FALSE}
rm(list = ls())
setwd("~/Documents/Baseline/baseline/data")
#setwd("C:/Users/bkram/CloudStation/HEART Lab/Baseline Study/baseline/data")
for (pkg in c("tidyverse", "cowplot", "plotly", "infer", "psych", "ggpubr", "grid", 
              "rcompanion", "nlme", "lme4", "lmerTest", "car", "multcompView", 
              "lsmeans", "gridExtra", "arse", "rstanarm", "bayestestR")) {library(pkg, character.only = TRUE)}
all_data <- read_csv("baseline_joined_data.csv") 
phys_data <- read.csv("baseline_physdata_091419.csv") %>% select(-1) %>% arrange(SubjectID) %>% drop_na()
```

```{r importing data, warning=FALSE}

combined_data <- all_data %>% 
  inner_join(phys_data, by = "SubjectID") %>% 
  select(SubjectID, HeartID, Condition, Semester, AppointmentTime, Segment, SegmentNumber, 
         MeanHeartRate, RSA, RMSSD, MeanIBI, RspRate, InitialSBP, InitialDBP, 
         BaselineStart, BaselineEnd, StroopStart, StroopEnd, PreBaselineStress, PostBaselineStress, 
         WorkStation, CNAPLabel, MP150Label, FingerCuffLabel, FingerCuffSize, ArmCuffLabel, ArmCuffSize, ECGLabel) %>% 
  mutate(Segment = dplyr::recode(Segment, A1 = "Baseline", A2 = "Stressor", A3 = "Recovery 1", A4 = "Recovery 2"))


all_data %>% group_by(Condition) %>% tally
all_data %>% group_by(AppointmentTime) %>% tally()
all_data %>% group_by(Semester) %>% tally()

# need to import bpv_data, survey_part1, survey_part3 and stroop_data next 
# need to recode date by semester and then control for research, semester, baseline starts and ends, etc.

```


```{r baseline variation in MeanHeartRate by participant, fig.width=9.5, fig.height=5}

colorblock_byppt <- combined_data %>% 
  filter(Condition == "Color Block") %>% 
  filter(SubjectID != "26") %>% # removing ppt 26 because of potential issues
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Color Block") +xlab("Baseline (in Minutes)") + ylab("Mean HR") 

magazines_byppt <- combined_data %>% 
  filter(Condition == "Magazines") %>% 
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Magazines") + xlab("Baseline (in Minutes)") + ylab("Mean HR") 

music_byppt <- combined_data %>% 
  filter(Condition == "Music") %>% 
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Music") + xlab("Baseline (in Minutes)") + ylab("Mean HR") 

questionnaires_byppt <- combined_data %>% 
  filter(Condition == "Questionnaires") %>% 
  filter(SubjectID != "61") %>% # removing ppt 61 because of potential issues
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Questionnaires") + xlab("Baseline (in Minutes)") + ylab("Mean HR")

resting_byppt <- combined_data %>% 
  filter(Condition == "Resting") %>% 
  filter(SubjectID != "302") %>% # removing ppt 302 because of potential issues
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Resting") + xlab("Baseline (in Minutes)") + ylab("Mean HR")

video_byppt <- combined_data %>% 
  filter(Condition == "Video") %>% 
  filter(Segment == "Baseline") %>% 
  ggplot(aes(x=SegmentNumber, y=MeanHeartRate, group=SubjectID)) + geom_line() +
  ggtitle("Video") + xlab("Baseline (in Minutes)") + ylab("Mean HR")

plot_grid(colorblock_byppt, magazines_byppt, music_byppt, questionnaires_byppt, resting_byppt, video_byppt, cols = 3)

ggplotly(colorblock_byppt) # possible issue with ppt 26
ggplotly(magazines_byppt) # looks fine 
ggplotly(music_byppt) # looks fine
ggplotly(questionnaires_byppt) # possible outlier on high side (ppt 61)
ggplotly(resting_byppt) # possible issue with 302 (major dip)
ggplotly(video_byppt) # looks fine

```

After visual inspection, it seems that there may be some potential outliers, including: 

- Color Block: Participant 26 (Segments 8-10)
- Questionnaires: Participant 4, 61 & 130 (Potential outliers)
- Resting: Participant 82 (Potential outlier) & Participant 302 (Segment 5) 

First, let's check if these outliers are actually outliers or not. The norm in psych seems to be exclude based on those outside of the +/-3 SDs. In past projects, we have done this on the overall baseline period, but I want to be a little more conservative with this project and check each minute. 

```{r}

descriptives <- describe(combined_data$MeanHeartRate)
descriptives$mean + (3 * descriptives$sd) # 108.0137
descriptives$mean - (3 * descriptives$sd) # 47.36618

```

Participant 61 is the only one that is consistently outside of the +/-3 SDs so we will remove them. I've also identified some issues with participants that 26 & 302 that are almost undoutedly HR cleaning errors in Mindware. Ideally, we will go back to check whether these data were due to coding problems in the HRV data rather than tossing these out. However, that hardware is in New Jersey so Brandon will have to do this when he gets access to the Mindware software again.

```{r mean variation over baseline, fig.width=9.5, fig.height=5}

colorblock_mbl <- combined_data %>% 
  filter(Condition == "Color Block") %>% 
  filter(Segment == "Baseline") %>% 
  filter(SubjectID != "26") %>%  # removing ppt 26 because of potential issue 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Color Block")) 
  
magazines_mbl <- combined_data %>% 
  filter(Condition == "Magazines") %>% 
  filter(Segment == "Baseline") %>% 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Magazines"))

music_mbl <- combined_data %>% 
  filter(Condition == "Music") %>% 
  filter(Segment == "Baseline") %>% 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Music"))

questionnaires_mbl <- combined_data %>% 
  filter(Condition == "Questionnaires") %>% 
  filter(Segment == "Baseline") %>% 
  filter(SubjectID != "61") %>% # removing ppt 61 because of potential issue 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Questionnaires"))

resting_mbl <- combined_data %>% 
  filter(Condition == "Resting") %>% 
  filter(Segment == "Baseline") %>% 
  filter(SubjectID != "302" ) %>%  # removing ppt 82 + 302 because of potential issue 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Resting"))

video_mbl <- combined_data %>% 
  filter(Condition == "Video") %>% 
  filter(Segment == "Baseline") %>% 
  group_by(SegmentNumber) %>% 
  summarise(heart_rate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Video"))

mbl <- bind_rows(colorblock_mbl, magazines_mbl, music_mbl, questionnaires_mbl, resting_mbl, video_mbl)  

mbl_graph <- ggplot(mbl, aes(x=SegmentNumber, y=heart_rate, group=Condition, color=Condition)) + geom_line(size = 1.3) +
  ggtitle("Mean Heart Rate Variation") + ylab("Mean Heart Rate") + xlab("Baseline (in Minutes)") + theme(legend.title = element_blank())

ggplotly(mbl_graph)

```

```{r mean variation over the entire experiment, fig.width=9.5, fig.height=5}

colorblock_byseg <- combined_data %>% 
  filter(Condition == "Color Block") %>%
  filter(SubjectID != "26") %>%  # removing ppt 26 because of potential issue
  group_by(Segment) %>% 
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Color Block")) %>% 
  select(Condition, Segment, MeanHeartRate) 

magazines_byseg <- combined_data %>% 
  filter(Condition == "Magazines") %>%
  group_by(Segment) %>% 
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Magazines")) %>% 
  select(Condition, Segment, MeanHeartRate) 

music_byseg <- combined_data %>% 
  filter(Condition == "Music") %>%
  group_by(Segment) %>% 
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Music")) %>% 
  select(Condition, Segment, MeanHeartRate)

questionnaires_byseg <- combined_data %>% 
  filter(Condition == "Questionnaires") %>%
  group_by(Segment) %>% 
  filter(SubjectID != "61") %>% # removing ppt 61 because of potential issue
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Questionnaires")) %>% 
  select(Condition, Segment, MeanHeartRate) 

resting_byseg <- combined_data %>% 
  filter(Condition == "Resting") %>%
  group_by(Segment) %>% 
  filter(SubjectID != "302" ) %>% # removing ppt 302 because of potential issue 
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>%
  mutate(Condition = paste("Resting")) %>% 
  select(Condition, Segment, MeanHeartRate) 

video_byseg <- combined_data %>% 
  filter(Condition == "Video") %>%
  group_by(Segment) %>% 
  summarise(MeanHeartRate = mean(MeanHeartRate)) %>% 
  mutate(Condition = paste("Video")) %>% 
  select(Condition, Segment, MeanHeartRate) 

byseg <- bind_rows(colorblock_byseg, magazines_byseg, music_byseg, questionnaires_byseg, resting_byseg, video_byseg)  

byseg_graph <- ggplot(byseg, aes(x=Segment, y=MeanHeartRate, group=Condition, color=Condition)) + 
  geom_line(size = 1.3) +
  ggtitle("Mean Heart Rate Variation") + ylab("Mean Heart Rate") + 
  theme(axis.title.x=element_blank()) + theme(legend.title = element_blank()); ggplotly(byseg_graph)

ggplotly(byseg_graph)

```

```{r anovas & ancovas}

combined_data$HeartID <- factor(combined_data$HeartID)
combined_data$Condition <- factor(combined_data$Condition)
combined_data <- combined_data %>% 
  mutate(Time = dplyr::recode(Segment, "Baseline" = 1, "Stressor" = 2, 
                              "Recovery 1" = 3, "Recovery 2" = 4))
combined_data

# within subject anova 
model1 <- aov(MeanHeartRate ~ Condition * Segment + Error(HeartID), data = combined_data)
summary(model1)

# in the past, we ran the anova > then did t-tests between different conditions > corrected with tukey analyses post hoc > but we need to run ancovas that control for baseline (and potentially other factors) and use the arse package to measure AUC (although i also saw some suggestions to use MLMs or a Bayesian approach to deal with time-series variations) > this is where i am now 

```

```{r}

test_min <- combined_data %>% 
  filter(Segment == "Stressor") %>% 
  group_by(HeartID) %>% 
  count()

min(test_min$n)

test_data <- combined_data %>%
  group_by(HeartID, Segment) %>% 
  select(HeartID, MeanHeartRate) %>% 
  mutate(testno = row_number())

new_test_data <- reshape(as.data.frame(test_data), idvar = "HeartID", timevar = "testno", direction = "wide") 

new_test_data
```



```{r}
combined_data %>% 
  group_by(Condition) 

combined_data$MeanHeartRate

combined_data %>% 
  mutate(HeartID = factor(combined_data$HeartID),
         Condition = factor(combined_data$Condition),
         Time = dplyr::recode(Segment, "Baseline" = 1, "Stressor" = 2, 
                              "Recovery 1" = 3, "Recovery 2" = 4)) %>% 
  group_by(Time) %>% 
  summarise(MeanHeartRate = as.numeric(mean(MeanHeartRate))) %>% 
  spread("MeanHeartRate")  ## cannot get spread to work 

?spread

#arse(data = test_df, xcoord = Time, ycoord = MeanHeartRate, saveout = TRUE)



xc <- t(c(1,2,3,4,5,6,7,8,9,10))
yc <- t(c(75,75,77,79,80,81,77,76,77,77))
dataset1 <- data.frame(xc, yc)
arse(data = dataset1, xcoord = dataset1[,1:10], ycoord = dataset1[,11:20], saveout = TRUE)
data

```




```{r other notes}



anova_data <- byseg

# ordering the factors for condition and segment 
anova_data$Condition <- factor(anova_data$Condition)
anova_data$Segment <- factor(anova_data$Segment, levels = c("Baseline", "Stressor", "Recovery 1", "Recovery 2")) 
str(anova_data$Segment)




model <- lm(MeanHeartRate ~ Condition + Segment + Condition * Segment, anova_data)
summary(model)
anova(model)

# Compact Model
baseline <- lme(MeanHeartRate ~ 1, 
                random = ~1 | Condition/Segment, 
                data = anova_data, 
                method = "ML")

# Augmented Model
nextModel <- lme(MeanHeartRate ~ Segment, 
                    random = ~1 | Condition/Segment, 
                    data = anova_data, 
                    method = "ML")

nextModel <- lme(MeanHeartRate ~ Segment + , 
                    random = ~1 | Condition/Segment, 
                    data = anova_data, 
                    method = "ML")

anova(baseline, nextModel)


byseg_graph <- ggplot(byseg, aes(x=Segment, y=MeanHeartRate, group=Condition, color=Condition)) + 
  geom_line(size = 1.3) +
  ggtitle("Mean Heart Rate Variation") + ylab("Mean Heart Rate") + 
  theme(axis.title.x=element_blank()) + theme(legend.title = element_blank()); ggplotly(byseg_graph)



```






```


References

https://rcompanion.org/handbook/I_09.html
http://www.jason-french.com/tutorials/repeatedmeasures.html
https://stats.idre.ucla.edu/r/seminars/repeated-measures-analysis-with-r/
https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf
http://www.tqmp.org/RegularArticles/vol15-2/p148/p148.pdf
https://cran.csiro.au/web/packages/arse/arse.pdf


bayesian
https://easystats.github.io/bayestestR/articles/bayestestR.html
https://easystats.github.io/bayestestR/articles/example1.html






























